Looking for a Signal in the Noise: Revisiting Obesity and the microbiome

### 2016-03-04 Sze_ObesityMicrobiomeMeta_PeerJ_2016.Rmd
manuscript update: abstract and Figs

**Results:** When we considered each dataset in isolation, there were no significant differences (P > 0.05) in B/F between non-obese and obese individuals. The Shannon diversity (N=1 study), OTU richness (N=2), and overall community dissimilarity (N=1) were significantly different between non-obese and obese individuals in at least one study (P<0.05). With individual studies, random forest was effective at identifying features within the datasets for differentiating between non-obese and obese individuals (mean AUC: 0.76; s.d.=0.09). In seven of the studies an OTU related to the Ruminococcaceae family was important in classification; however, the direction of the association varied between studies. When we pooled the results across studies, there was a significant difference in RR of obesity based on Shannon diversity (RR = 1.19, CI = 1.04-1.36, P-value = 0.01) and B/F ratio (RR = 0.86, CI = 0.75-0.99, P-value = 0.03). Finally, when we normalized the Shannon diversity data and pooled studies there was a significant difference between non-obese and obese individuals (difference in z-score=0.XX; P=0.0X) something about... B/F.

@@

We attempted to stick to the original PRISMA guidelines for performing a meta-analysis as published by Moher et al [-@moher_preferred_2010]. A detailed description of the literature seach and the exact formula we used for the initial PubMed search can be found in the online supplement. In brief, a total of 196 studies made it through the initial identification of studies. Screening removed a total of 184 studies leaving a total of 12 eligible studies for this meta-analysis. Upon a full reading of these specific manuscripts 2 were exculded and the reasons can be found in the online supplement. A total of 10 studies were ultimately deemed eligible for qualitative synthesis [Figure 1]. One final study was removed from the quantitative analysis since no individual in the study was found to be able to be classified in the obese category based on the classical cutoffs. A total of 3 studies were from records identified by database searching [@turnbaugh_core_2009; @zupancic_analysis_2012; @escobar_gut_2014] and the other 6 were from records identified by other sources [@ross_16s_2015; @arumugam_enterotypes_2011; @goodrich_human_2014; @wu_linking_2011; @conlan_species-level_2012]. One of these studies (Baxter) has not been published at the time of this writing.

@@

**Figure 5: Pooled Analysis of Shannon Diversity and B/F ratio by either Obesity Status or BMI Groups.**  **A)** ZScore Normalized Shannon Diversity and Obesity. There was a significant difference between non-obese (n=1099) and obese groups (n=435) (P-value = 0.005992). **B)** ZScore Normalized Shannon Diversity by BMI Group. Total n for lean, overweight, and obese groups are 575, 435, 524 respectively. Using an ANOVA with tukey post-hoc testing there was a significant diference between the overweight and obese group (P-value = 0.018) but no difference between lean and obese group (P-value = 0.081). **C)** ZScore Normalized B/F Ratio and Obesity. There was no significant difference between the non-obese (n=1099) and obese (n=435) group based on B/F ratio (P-value = 0.456). **D)** ZScore Normalized B/F Ratio by BMI Group. Total n for lean, overweight, and obese groups are 575, 435, 524 respectively. There was no significant difference between any of the three groups based on ANOVA with Tukey post-hoc testing (P-value = 0.398).

>>>>

**Results:** When we considered each dataset in isolation, there were no significant differences (P > 0.05) in B/F between non-obese and obese individuals. The Shannon diversity (N=1 study), OTU richness (N=2), and overall community dissimilarity (N=1) were significantly different between non-obese and obese individuals in at least one study (P<0.05). With individual studies, random forest was effective at identifying features within the datasets for differentiating between non-obese and obese individuals (mean AUC: 0.76; s.d.=0.09). In seven of the studies an OTU related to the Ruminococcaceae family was important in classification; however, the direction of the association varied between studies. When we pooled the results across studies, there was a significant difference in RR of obesity based on Shannon diversity (RR = 1.19, CI = 1.04-1.36, P-value = 0.02) and B/F ratio (RR = 0.86, CI = 0.75-0.99, P-value = 0.04). Finally, when we normalized the Shannon diversity data and pooled studies there was a significant difference between non-obese and obese individuals (difference in z-score=0.155; P=0.006) something about... B/F.

@@

We attempted to stick to the original PRISMA guidelines for performing a meta-analysis as published by Moher et al [-@moher_preferred_2010]. A detailed description of the literature seach and the exact formula we used for the initial PubMed search can be found in the online supplement. In brief, a total of 196 studies made it through the initial identification of studies. Screening removed a total of 184 studies leaving a total of 12 eligible studies for this meta-analysis. Upon a full reading of these specific manuscripts 2 were exculded and the reasons can be found in the online supplement. A total of 10 studies were ultimately deemed eligible for qualitative synthesis [Figure 1]. One final study was removed from the quantitative analysis since no individual in the study was found to be able to be classified in the obese category based on the classical cutoffs. A total of 3 studies were from records identified by database searching [@turnbaugh_core_2009; @zupancic_analysis_2012; @escobar_gut_2014] and the other 6 were from records identified by other sources [@ross_16s_2015; @arumugam_enterotypes_2011; @goodrich_human_2014; @wu_linking_2011; @conlan_species-level_2012]. One of these studies (Baxter) was used on a data set that was not originally designed to investigate obesity in order to help mitigate any possible publication bias associated with the bacterial microbiome and obesity.

@@

**Figure 5: Pooled Analysis of Shannon Diversity and B/F ratio by either Obesity Status or BMI Groups.**  **A)** ZScore Normalized Shannon Diversity and Obesity. There was a significant difference between non-obese (n=1081) and obese groups (n=435) (P-value = 0.005992). **B)** ZScore Normalized Shannon Diversity by BMI Group. Total n for lean, overweight, and obese groups are 563, 435, 518 respectively. Using an ANOVA with tukey post-hoc testing there was a significant diference between the overweight and obese group (P-value = 0.018) but no difference between lean and obese group (P-value = 0.081). **C)** ZScore Normalized B/F Ratio and Obesity. There was no significant difference between the non-obese (n=1081) and obese (n=435) group based on B/F ratio (P-value = 0.456). **D)** ZScore Normalized B/F Ratio by BMI Group. Total n for lean, overweight, and obese groups are 563, 435, 518 respectively. There was no significant difference between any of the three groups based on ANOVA with Tukey post-hoc testing (P-value = 0.398).
